---
title: "Tidy StatQuest - 01"
author: "Jonathan Kitt"
date: "2026-07-03"
draft: true
categories: [StatQuest, tidy]
---

A few years ago, I started watching Joshua Starmer's Youtube videos.

When his first book came out, I quickly got my hands on a copy, and I'm now the happy owner of his StatQuest trilogy.

Josh's videos and books have been a huge help as I embarked on my statistics/data science journey.

For each chapter of his books, Josh wrote companion coding tutorials, available on Github.

In this series, I'll be "translating" the coding tutorials into {tidyverse} syntax.

Let's get started!

| Book | Chapter | Coding tutorial |
|----|----|----|
| The StatQuest Illustrated Guide to Statistics | 01 - Fundamental Concepts in Statistics!!! | <https://github.com/StatQuest/sigs/tree/main/chapter_01> |

# Set up

```{r}
# Load packages 
library(tidyverse)
library(janitor)
library(rstatix)
```

# Load data from a file

The data is available on the Github repository. We download it using the {readr} package:

```{r}
file_url <- "https://raw.githubusercontent.com/StatQuest/sigs/refs/heads/main/chapter_01/spend_n_save.txt"

spend_n_save <- read_tsv(file_url) |> 
  clean_names() |> 
  mutate(id = factor(id))
```

Print out the first and last rows:

```{r}
# Print out the first 5 rows
head(spend_n_save, n = 5)

# Print out the last 5 rows
tail(spend_n_save, n = 5)
```

# Population Mean and Standard Deviation

Calculate the mean and standard deviation of the number of apples:

```{r}
pop_stats <- spend_n_save |> 
  get_summary_stats(type = "mean_sd")

pop_stats
```

# Estimated Mean and Standard Deviation

```{r}
set.seed(42)

est_stats <- spend_n_save |> 
  slice_sample(n = 5) |> 
  get_summary_stats(type = "mean_sd")

est_stats
```

Calculate the biased standard deviation for the sample:

```{r}
est_stats |> 
  rename(sd_unbiased = sd) |> 
  mutate(sd_biased = round(sqrt(sd_unbiased^2 * (n - 1) / n), digits = 3))
```

# Simulate biased and unbiased estimates

```{r}
set.seed(42)

var_unbiased <- replicate(n = 1000, expr = {
  spend_n_save |> 
  slice_sample(n = 5) |> 
  summarise(var_unbiased = var(num_apples)) |> 
  pull()
})
```

```{r}
tibble(
  n = 5,
  var_unbiased = var_unbiased
) |> 
  mutate(sd_unbiased = sqrt(var_unbiased),
         var_biased = var_unbiased * (n - 1) / n,
         sd_biased = sqrt(var_biased)
         ) |> 
  summarise(mean_sd_unbiased = round(mean(sd_unbiased), 1),
            mean_sd_biased = round(mean(sd_biased), 1))
```
